---
title: "ml_workflow"
author: "WangYong"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Target
The goal of this competition is to predict rainfall for each day of the year.

ML tools: 
  tidymodels related worksflows & glm, lightgbm, ranger engine.
Evaluatioin metric:
  roc_auc, percent of rain probability.
  sample_submission.csv : 0.5. it is value is 0. I tried change it to 1, the score is 0.5 as well.
  kaggle best score is 0.87 in the begining.

## librar y & load_data

### library

```{r}
library(tidyverse)
library(tidymodels)
library(finetune)
library(future)
library(purrr)
library(furrr)
library(textrecipes)
library(themis)


library(bonsai)
library(lightgbm)
library(xgboost)
library(ranger)

library(readr)
library(janitor)
library(lubridate)
```

```{r}
#TODO, column 18,7 has some strange data leave it handle future. 71 lines in train, 159 line in test
# row833049 column 18 is f has-ring is total wrong. 
# so keep using readr:read_csv handling
```

### loading data

```{r}
data_path <- '../input/playground-series-s5e3/'
train<- 
  readr::read_csv(file.path(data_path, 'train.csv'),
                  show_col_types = FALSE)|>
  mutate(rainfall=as.factor(rainfall))
test <- 
  readr::read_csv(file.path(data_path, 'test.csv'),
                  show_col_types = FALSE)
submission <-  readr::read_csv(file.path(data_path, 'sample_submission.csv'),show_col_types = FALSE)
```

### quick skim

```{r}
train|> skimr::skim()
```

```{r}
test|> skimr::skim()
```

```{r}
submission |> skimr::skim()
```

### check if train & test is same distribution

```{r}
get_df_var<-function(df){
  df|>
    select(-any_of(c('id','rainfall')))|>
    summarize_all(var)|>
    pivot_longer(cols=everything(),
                 names_to='feature',
                 values_to='variance')

}
list(train=train, test=test)|>
  map_dfr(\(x) get_df_var(x), .id = "dataset") |>
  pivot_wider(names_from=dataset, values_from = variance)|>
  mutate(pct_change=(train-test)/train)#|>arrange(desc(abs(diff)))
```

### Finding of different distribution
there is no big change found 

## EDA

### time series EDA
```{r}
library(fpp3)

trans_tsibble <-function(){

  tmp <- bind_rows(
    list(train=train,test=test),
    .id = 'source') |>
    mutate(
      source=as.factor(source),
      day=case_when(id==1037~307+1,
                    id == 1132~37+1,
                    id ==1210~115+1,
                    id ==1132~265+1,
                    id ==1251~156+1,
                    id ==1284~189+1,
                    id == 1290~195+1,
                    id== 1312~217+1,
                    id == 1318~223+1,
                    id ==1346~251+1,
                    id == 1352~257+1,
                    id == 1367~272+1,
                    id ==  1373~278+1,
                    id ==1380~285+1,
                    id == 1382~287+1,
                    id ==1388~293+1,
                    id == 1395~300+1,
                    id == 1400~305+1,
                    id == 1403~308+1,
                    id ==1406~311+1,
                    id == 1404~309+1,
                    id == 1407~312+1,
                    id == 1409~314+1,
                    id == 1414~319+1,
                    id == 1416~321+1,
                    id ==1420 ~325+1,
                    id == 1428~333+1,
                    id ==1430~335+1,
                    id ==1438~343+1,
                    id == 1439~344+1,
                    id == 1445 ~ 350+1,
                    id == 1452 ~ 357 +1,
                    id == 1453 ~ 358+1,
                    id == 1457 ~ 362+1,
                    id == 1458 ~ 363+1,
                    id == 1459 ~ 364+1,
                    .default=day),
      year_offset = case_when( id >=0 & id <365~1,
                               id >=365 & id <730~2,
                               id >=730 & id <1095~3,
                               id >=1095 & id <1460~4,
                               id >=1460 & id <1825~5,
                               id >=1825 & id <2190~6,
                               id >=2190 & id <2555~7,
                               id >=2555 ~8))|>
    mutate(date = make_date(year = 2020 + year_offset, # create the year column for the date
                            month = 1,
                            day =1) + days(day -1) # days is offset from teh start of the year
    )
  tmp|>
    as_tsibble(index=date)
}

ts_df <- trans_tsibble()
ts_df |> ggplot2::ggplot(aes(x=id, y=day,color=source)) + geom_point()


```
#### plot ts
```{r}
ts_df |>as_tibble()|>
  select(-any_of(c('id','day', 'rainfall','year_offset','date')))|>
  group_by(source)|>
  summarize_all(var)|>
  pivot_longer(-source)|>
  pivot_wider(names_from = source, values_from=value)|>
  mutate(pct_change= (train-test)/train)
```



## coding

### 1. Data Loading and Initial Exploration ----

### 2. Feature Engineering ----

-   leave it in the preprocessing recipe

### 3. Data Splitting ----

#### augment_df

#### split/cv

```{r}
set.seed(1234)
train <- ts_df |> as_tibble() |>filter(source=='train')|>select(-source)
df_split <- initial_time_split(train, prop = 0.8)
 train_set <- training(df_split)
 test_set <- testing(df_split)
#cv_folds <- vfold_cv(train_set,v = 10,strata=rainfall)
cv_folds <- train_set |> sliding_period(index=date, period='year')
```

### 4. Preprocessing Recipe ----

#### 4.0 v0 base_line - 
```{r}
rcp_bs_v0 <-
  recipe(rainfall ~ ., data = train_set) |>
  update_role(id, new_role='ID')|>
  step_rm(date,year_offset)|>
  step_impute_median(all_numeric_predictors())|> 
  step_log(all_numeric_predictors(),offset = 1, skip = FALSE) |>
  # process the logical feature to factor feature
  step_bin2factor(all_logical_predictors())|>
  # proceeding the convert the character feature to factorical feature
  step_novel(all_nominal_predictors())|>
  step_unknown(all_nominal_predictors()) |>
  step_other(all_nominal_predictors())|>
  step_dummy(all_nominal_predictors(),one_hot = TRUE) |>
  step_nzv(all_predictors())|>
  #step_corr(all_numeric_predictors())|>
  step_normalize(all_numeric_predictors())|> # Scale numeric predictors
  #step_smote(rainfall,over_ratio = 1,skip = TRUE)|>
  check_missing(all_predictors())

```

#### 4.1 v1 base_line - bad
```{r}
rcp_bs_v1 <-
  recipe(rainfall ~ ., data = train_set) |>
  update_role(id, new_role='ID')|>
  #step_date(date, features=c('month','week'))|>
   step_mutate(
   temp_range= maxtemp - mintemp,
   heat_index= temparature+ 0.5 * (temparature- 10) * (humidity / 100),
   dew_dep= temparature - dewpoint,
   cloud_sun_ratio = cloud / sunshine + 1,    
   pressure_change = c(NA, diff(pressure)),
   air_density = 1.225 , 
   wind_power = 0.5 * air_density * windspeed**3,
   wind_direction_cat = cut(winddirection, 
                            breaks=c(0, 90, 180, 270, 360), 
                            labels=c("North", "East", "South", "West"), 
                            include.lowest=TRUE),
   cloud_humidity = cloud + humidity,
   cloud_humidity_sunshine = cloud + humidity + sunshine,
   cloud_x_sunshine = cloud * sunshine,
   humidity_x_sunshine = humidity * sunshine)|>
  step_rm(date,year_offset)|>
  step_impute_median(all_numeric_predictors())|> 
  step_log(all_numeric_predictors(),offset = 1, skip = FALSE) |>
  step_lag(all_numeric_predictors(),lag =c(1))|>
  step_impute_median(all_numeric_predictors())|> 
  # process the logical feature to factor feature
  step_bin2factor(all_logical_predictors())|>
  # proceeding the convert the character feature to factorical feature
  step_novel(all_nominal_predictors())|>
  step_unknown(all_nominal_predictors()) |>
  step_other(all_nominal_predictors())|>
  step_dummy(all_nominal_predictors(),one_hot = TRUE) |>
  step_nzv(all_predictors())|>
  step_corr(all_numeric_predictors())|>
  step_normalize(all_numeric_predictors())|> # Scale numeric predictors

  check_missing(all_predictors())
rcp_bs_v1 |>prep()|juice()
```

#### 4.2 v2 base_line - 
```{r}
# df["Temp_Range"] = df["maxtemp"] - df["mintemp"]
#     df["Heat_Index"] = df["temparature"] + 0.5 * (df["temparature"] - 10) * (df["humidity"] / 100)    
#     df["Dew_Dep"] = df["temparature"] - df["dewpoint"]    
#     df["Wind_Chill"] = np.where(
#         df["temparature"] < 10,
#         13.12 + 0.6215 * df["temparature"] - 11.37 * df["windspeed"]**0.16 + 
#        0.3965 * df["temparature"] * df["windspeed"]**0.16,
#         df["temparature"]
#     )    
#     df["Cloud_Sun_Ratio"] = df["cloud"] / (df["sunshine"] + 1)    
#     df["Pressure_Change"] = df["pressure"].diff().fillna(0)    
#     air_density = 1.225  
#     df["Wind_Power"] = 0.5 * air_density * df["windspeed"]**3         
#     df["Wind_Direction_Cat"] = pd.cut(df["winddirection"], 
#                                       bins=[0, 90, 180, 270, 360], 
#                                       labels=["North", "East", "South", "West"], 
#                                       include_lowest=True)
# 
#     df['cloud + humidity'] = df.cloud + df.humidity
#     df['cloud + humidity + sunshine'] = df.cloud + df.humidity + df.sunshine
#     df['cloud * sunshine'] = df.cloud * df.sunshine
#     df['humidity * sunshine'] = df.humidity * df.sunshine
# 
#     df["HTI"] = df["temparature"] + 0.2 * df["humidity"]    
#     df["ACI"] = (df["maxtemp"] + df["mintemp"]) / 2 - (df["humidity"] / 2) + df["dewpoint"]    
#     df["CSI"] = (df["sunshine"] - df["cloud"]) / (df["sunshine"] + df["cloud"] + 1)  
#     df["WCI"] = df["temparature"] - 0.5 * df["windspeed"]
#     df['day'] = pd.to_datetime(df['day'], errors='coerce')    
#    
#     df['month'] = df['day'].dt.month
#     df['day_of_week'] = df['day'].dt.dayofweek
#     df['is_weekend'] = df['day_of_week'].isin([5, 6]).astype(int)    
#     df['temp_range'] = df['maxtemp'] - df['mintemp']
#     df['avg_temp'] = (df['maxtemp'] + df['mintemp']) / 2
#     df['temp_deviation'] = df['temparature'] - df['avg_temp']   
#   
#     df['dew_point_depression'] = df['temparature'] - df['dewpoint']    
#    
#     df['wind_dir_rad'] = np.deg2rad(df['winddirection'])
#     df['wind_dir_sin'] = np.sin(df['wind_dir_rad'])
#     df['wind_dir_cos'] = np.cos(df['wind_dir_rad'])
#     df.drop(columns=['wind_dir_rad'], inplace=True)   
#     df['wind_chill'] = 13.12 + 0.6215 * df['temparature'] - 11.37 * (df['windspeed']**0.16) + 0.3965 * df['temparature'] * (df['windspeed']**0.16)
#     df['humidity_temp'] = df['humidity'] * df['temparature']
#     df['cloud_sunshine'] = df['cloud'] * df['sunshine']
#     df['rolling_temp_mean'] = df['avg_temp'].rolling(window=7).mean()
#     df['rolling_wind_mean'] = df['windspeed'].rolling(window=7).mean()
#     df['rolling_humidity_mean'] = df['humidity'].rolling(window=7).mean()
#     df['temp_lag_1'] = df['avg_temp'].shift(1)
#     df['humidity_lag_1'] = df['humidity'].shift(1)
#     df['windspeed_lag_1'] = df['windspeed'].shift(1)
#     df['pressure_temp_interaction'] = df['pressure'] * df['avg_temp']
#     df['windspeed_temp_interaction'] = df['windspeed'] * df['avg_temp']
#     df['sunshine_cloud_interaction'] = df['sunshine'] * df['cloud']
#     df['season'] = df['month'].apply(lambda x: 'Spring' if 3 <= x <= 5 else
#                                       'Summer' if 6 <= x <= 8 else
#                                       'Autumn' if 9 <= x <= 11 else 'Winter')
#     for c in ['pressure', 'maxtemp', 'temparature', 'humidity']:
#         for gap in [1]:
#             df[c+f"_shift{gap}"] = df[c].shift(gap)
#             df[c+f"_diff{gap}"] = df[c].diff(gap)
#     df = pd.get_dummies(df, columns=['season'], drop_first=True)  
#     df.drop(columns=['day'], inplace=True
```

```{r}

rcp_bs_v2 <-
  recipe(rainfall ~ ., data = train_set) |>
  update_role(id, new_role='ID')|>
  step_rm(date,year_offset)|>
  step_impute_median(all_numeric_predictors())|> 
  step_mutate(
   temp_range= maxtemp - mintemp,
   heat_index= temparature+ 0.5 * (temparature- 10) * (humidity / 100),
   dew_dep= temparature - dewpoint,
   cloud_sun_ratio = cloud / sunshine + 1,    
   pressure_change = c(NA,diff(pressure)),
   air_density = 1.225 , 
   wind_power = 0.5 * air_density * windspeed**3,
   wind_direction_cat = cut(winddirection, 
                            breaks=c(0, 90, 180, 270, 360), 
                            labels=c("North", "East", "South", "West"), 
                            include.lowest=True),
   cloud_humidity = cloud + humidity,
   cloud_humidity_sunshine = cloud + humidity + sunshine,
   cloud_x_sunshine = cloud * sunshine,
   humidity_x_sunshine = humidity * sunshine)|>
  step_impute_median(all_numeric_predictors())|> 
  step_log(all_numeric_predictors(),offset = 1, skip = FALSE) |>
  # process the logical feature to factor feature
  step_bin2factor(all_logical_predictors())|>
  # proceeding the convert the character feature to factorical feature
  step_novel(all_nominal_predictors())|>
  step_unknown(all_nominal_predictors()) |>
  step_other(all_nominal_predictors())|>
  step_dummy(all_nominal_predictors(),one_hot = TRUE) |>
  step_nzv(all_predictors())|>
  #step_corr(all_numeric_predictors())|>
  step_normalize(all_numeric_predictors())|> # Scale numeric predictors
  #step_smote(rainfall,over_ratio = 1,skip = TRUE)|>
  check_missing(all_predictors())

```

#### 4.10 all recipes

```{r}
set.seed(1234)
library(future)
library(furrr)
selected_rcps <- list(bs_v0=rcp_bs_v0,
                      bs_v1=rcp_bs_v1#, # is significanttly change the data , and performance is descress . give it up.
                      #bs_v2=rcp_bs_v2
                      )
plan(multisession,workers = 5)
#selected_rcps|>map(\(rcp_item) rcp_item|>prep()|>bake(new_data=train)|>summary())
plan(sequential)
```

### 5. Model Specification ----

```{r}
glm_eng <- 
  logistic_reg(penalty = 0.05,
               mixture = 0.05) |>  # Example penalty and mixture values
  set_engine("glmnet") |>
  set_mode("classification")    # Specify classification

lgbm_eng<-
   parsnip::boost_tree(
      trees = 700, # Number of trees
      learn_rate = 0.01,
      #tree_depth =6,
      loss_reduction = 0.001,
      #sample_size = 0.75, # Added sample_size
      #tree_depth = tune(),
      
      #mtry = tune(),
      #min_n = 20
   ) |>
   set_mode("classification")|>
   set_engine("lightgbm",
              #metric='roc_auc', 
              #num_leaves = 31,
              num_threads = 4,
              verbose=1) 

rf_eng<- rand_forest( trees = 700, 
                      #learn_rate = 0.01,
                      #mtry=tune(), 
                      min_n=30) |>
  set_engine("ranger",num.threads=4)|>
  set_mode("classification") 

xgb_eng<- parsnip::boost_tree( trees = 700, 
                      learn_rate = 0.01,
                      loss_reduction = 0.001,
                      sample_size = 0.8, # Added sample_size
                      #mtry=tune(),
                      min_n=20) |>
  set_engine("xgboost",num.threads=8)|>
  set_mode("classification")
selected_eng <- list(glm=glm_eng,
                     rf=rf_eng,
                     lgbm=lgbm_eng,
                     xgb=xgb_eng
                     )

```

### 6. Workflow ----
#### set metrics
```{r}
rocauc_metrics <- metric_set(roc_auc) # main goal is roc_auc, accuracy is just for reference
```

#### simple wflow

```{r}
set.seed(1234)
#plan(multisession, workers = 1)  
#selected_rcps |>
#furrr::future_map_dfr(\(rcp_item) 
simple_wf_fit <- 
  workflow() |>
  add_recipe(rcp_bs_v1) |>
  add_model(glm_eng)|>
  last_fit(df_split,
           metrics=rocauc_metrics)
#plan(sequential)
 simple_wf_fit |> collect_metrics()
#   extract_fit_engine()|>
#   plot()

```

#### simple workflowset

```{r}
set.seed(1234)
library(future)
plan(multisession,workers = 12)
ctrl <- control_resamples(save_pred = TRUE, save_workflow = TRUE,verbose=TRUE)
wfs_result <-
  workflow_set(preproc = selected_rcps,
               models = selected_eng ) |>
  workflow_map(fn='fit_resamples',
               resamples =cv_folds,
               metrics =rocauc_metrics,
               control = ctrl
               )
wfs_result|>collect_metrics()  |> filter(.metric=='roc_auc')
  
plan(sequential)
```

### 7 stacking

```{r}
set.seed(1234)
library(future)
#plan(multisession,workers = 4)
combined_fit <-
  stacks::stacks()|>
  stacks::add_candidates(wfs_result)|>
  stacks::blend_predictions()|>
  stacks::fit_members()

combined_fit|>
  autoplot(type = "weights")

autoplot(combined_fit)
#plan(sequential)
```

### 7. Tuning Grid ----

```{r}
tune_lgbm_spec <-
  boost_tree(
    trees = tune(),             # Number of trees (boosting rounds)
    mtry = tune(),              # Number of features to consider at each split
    learn_rate = tune(),        # Learning rate (shrinkage)
    tree_depth = tune(),         # Maximum depth of each tree
    min_n = tune(),            # Minimum number of data points in a node
    loss_reduction = tune()      # Complexity parameter
  ) %>%
  set_engine("lightgbm",
             num_leave=tune()) %>%
  set_mode("classification")

tune_lgbm_grid <- grid_space_filling(
  trees(range = c(500, 1000)),  # Wider range
  #mtry = mtry(range = c(0.5, 0.9)),    # Wider range
  learn_rate(range = c(-3, -1)),  # Log scale
  tree_depth(range = c(5, 12)),  # Wider range
  min_n(range = c(10, 60)),   # Wider range
  loss_reduction(range = c(0, 0.1)), # Wider range
  num_leaves(range=c(31,100)),
  size = 50 # Increased number of combinations to try
)

tune_lgbm_workflow <- 
  workflow_set(preproc = selected_rcps,
               models = list(lgb=lgbm_eng) ) 

# Tune the model using ANOVA Race
tune_results <-  tune_lgbm_workflow |>
  workflow_map(fn='tune_race_anova',
               resamples=cv_folds,
               seed=1234,
               grid=tune_lgbm_grid,
               metrics=rocauc_metrics,
               control = control_race(save_pred = TRUE, save_workflow = TRUE,verbose=TRUE) # Show progress
               )
tune_results|>rank_results()

```

### 8. Cross-Validation ----

```{r}
# combined it with step3 data splitting
```

### 9. Tuning and Evaluation ----

```{r}
# plan(multisession,workers =2)
# cars_tune_results <- cars_workflow |>
#   tune_grid(
#     resamples = cars_folds,
#     grid = cars_grid,
#     metrics = metric_set(rmse),
#      control = control_grid(save_pred = TRUE, 
#                             verbose = TRUE,
#                             allow_par = F) # Keep predictions
#   )
#  
#  # Find best parameters
#  best_params <- cars_tune_results |>
#    select_best("rmse")
# 
#  # Finalize workflow with best parameters
#  final_workflow <- cars_workflow |>
#    finalize_workflow(best_params)
```

```{r}
# Fit the final workflow to the training data
# final_lgbm_fit <- last_fit(final_workflow,cars_split )
# final_lgbm_mod <- extract_workflow(final_lgbm_fit )
# collect_metrics(final_lgmb_mod)

# plan(sequential)

```

### 10. Evaluate on Test Set ----

```{r}
combined_test_result <- 
  test_set %>%
  bind_cols(predict(combined_fit, 
                    new_data=test_set,type='prob'))
combined_test_result|>roc_auc(rainfall, .pred_1,event_level = 'second')
```

### 11. Prepare Submission ----

```{r}
set.seed(1234)
library(future)
plan(multisession,workers = 12)
#final_model <- combined_fit#simple_wf_fit|>extract_workflow()
final_model <- simple_wf_fit|>extract_workflow()
final_predictions <- final_model |>
   predict(new_data = ts_df|>as_tibble()|>filter(source=='test'),
           type='prob') 
plan(sequential)

 # #Handle negative predictions
 # final_predictions <- final_predictions |>
 #   mutate(.pred= ifelse(.pred< 0, 0, .pred))

 # Save submission file
 submission |>
   mutate(rainfall=final_predictions$.pred_1)|>
   readr::write_csv("submission.csv")
 zip('submission.csv.zip','submission.csv')
 
```

## kaggle submission



### score submit
```{r}
# submit latest submission.csv
system('kaggle competitions submit -c playground-series-s5e3 -f submission.csv.zip -m "sample_submission"')
Sys.sleep(15)
# get latest score 
system('kaggle competitions submissions -q -c playground-series-s5e3')
# 
# # get leader board score
# #system('kaggle competitions leaderboard -s -v -c playground-series-s4e8')
```

### notebook convert
```{r}
library(rmd2jupyter)
rmd2jupyter('rainfall_roc_auc_workflow.Rmd')
```
